
### Two types of standardization

- Direct standardization
- Indirect standardization

- Reference population is the `world_standard_population_by_sex`.

- Based on the populations of 46 countries

## load packages and datasets

```{r}
pacman::p_load(
     rio,                 # import/export data
     here,                # locate files
     stringr,             # cleaning characters and strings
     frailtypack,         # needed for dsr, for frailty models
     dsr,                 # standardise rates
     PHEindicatormethods, # alternative for rate standardisation
     tidyverse)           # data management and visualization
```

```{r}
# install the dsr package if you are using newer version of R
packageurl <- "https://cran.r-project.org/src/contrib/Archive/dsr/dsr_0.2.2.tar.gz"
install.packages(packageurl, repos=NULL, type="source")
```

```{r}
# import demographics for country A directly from Github
A_demo <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/standardization/country_demographics.csv") |> mutate(Country = "A") |> 
       mutate(
          age_cat5 = str_replace_all(age_cat5, "\\+", ""))

# import deaths for country A directly from Github
A_deaths <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/standardization/deaths_countryA.csv")


# import demographics for country B directly from Github
B_demo <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/standardization/country_demographics_2.csv") |> mutate(Country = "B")

# import deaths for country B directly from Github
B_deaths <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/standardization/deaths_countryB.csv")

# import demographics for standard population directly from Github
standard_pop_data <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/standardization/world_standard_population_by_sex.csv")
```

```{r}
# Country A
A_demo
B_demo
A_deaths
B_deaths
standard_pop_data
```
```{r}
# combine pop datasets
pop_countries <- A_demo %>%  # begin with country A dataset
     bind_rows(B_demo) %>%        # bind rows, because cols are identically named
     pivot_longer(                       # pivot longer
          cols = c(m, f),                   # columns to combine into one
          names_to = "Sex",                 # name for new column containing the category ("m" or "f") 
          values_to = "Population") %>%     # name for new column containing the numeric values pivoted
     mutate(Sex = recode(Sex,            # re-code values for clarity
          "m" = "Male",
          "f" = "Female"))
pop_countries
```

```{r}
# combine death datasets
deaths_countries <- A_deaths %>%    # begin with country A deaths dataset
     bind_rows(B_deaths) %>%        # bind rows with B dataset, because cols are identically named
     pivot_longer(                  # pivot longer
          cols = c(Male, Female),        # column to transform into one
          names_to = "Sex",              # name for new column containing the category ("m" or "f") 
          values_to = "Deaths") %>%      # name for new column containing the numeric values pivoted
     rename(age_cat5 = AgeCat)      # rename for clarity
deaths_countries
```

```{r}
# join pop and deaths datasets
country_data <- pop_countries %>% 
     left_join(deaths_countries, by = c("Country", "age_cat5", "Sex"))
country_data
```

```{r}
# reorder the country, sex and age category levels
country_data <- country_data %>% 
  mutate(
    Country = fct_relevel(Country, "A", "B"),
      
    Sex = fct_relevel(Sex, "Male", "Female"),
        
    age_cat5 = fct_relevel(
      age_cat5,
      "0-4", "5-9", "10-14", "15-19",
      "20-24", "25-29",  "30-34", "35-39",
      "40-44", "45-49", "50-54", "55-59",
      "60-64", "65-69", "70-74",
      "75-79", "80-84", "85")) %>% 
          
  arrange(Country, age_cat5, Sex)
country_data

```

```{r}
# clean reference population
# Remove specific string from column values
standard_pop_clean <- standard_pop_data %>%
  rename(age_cat5 = AgeGroup) %>%
     mutate(
          age_cat5 = str_replace_all(age_cat5, "years", ""),   # remove "year"
          age_cat5 = str_replace_all(age_cat5, "plus", ""),    # remove "plus"
          age_cat5 = str_replace_all(age_cat5, " ", "")) %>%   # remove " " space
     
     rename(pop = WorldStandardPopulation)   # change col name to "pop", as this is expected by dsr package
standard_pop_clean
```

```{r}
# join dataset standard population with the country data
all_data <- left_join(country_data, standard_pop_clean, by=c("age_cat5", "Sex"))
all_data
```

## Using the dsr package

- can only calculate direct standardized rates
- `dsr()` expects one data frame with the country and events population and another dataframe with the reference population. 
- It also expects that the reference population dataset that the population variable is represented as `pop`.


```{r}
# Calculate rates per country directly standardized for age and sex
mortality_rate <- dsr::dsr(
     data = country_data,  # specify object containing number of deaths per stratum
     event = Deaths,       # column containing number of deaths per stratum 
     fu = Population,      # column containing number of population per stratum
     subgroup = Country,   # units we would like to compare
     age_cat5,             # other columns - rates will be standardized by these
     Sex,
     refdata = standard_pop_clean, # reference population data frame, with column called pop
     method = "gamma",      # method to calculate 95% CI
     sig = 0.95,            # significance level
     mp = 100000,           # we want rates per 100.000 population
     decimals = 2)          # number of decimals)


mortality_rate
```
The standardized mortality rate is 1.22 times higher in country A compared to country B 95% CI(1.17 - 1.27)


### Standardized rate difference

```{r}
# Calculate RD
mortality_rd <- dsr::dsrr(
     data = country_data,       # specify object containing number of deaths per stratum
     event = Deaths,            # column containing number of deaths per stratum 
     fu = Population,           # column containing number of population per stratum
     subgroup = Country,        # units we would like to compare
     age_cat5,                  # characteristics to which we would like to standardize
     Sex,                        
     refdata = standard_pop_clean, # reference population, with numbers in column called pop
     refgroup = "B",            # reference for comparison
     estimate = "difference",   # type of estimate
     sig = 0.95,                # significance level
     mp = 100000,               # we want rates per 100.000 population
     decimals = 2)              # number of decimals

mortality_rd
```

Country A has 4.24 additional deaths per 100.000 population (95% CI 3.24-5.24) compared to country A.




## PHEindicatormethods pacakge

### Direct standardized rates

```{r}
# Calculate rates per country directly standardized for age and sex
mortality_ds_rate_phe <- all_data %>%
     group_by(Country) %>%
     PHEindicatormethods::phe_dsr(
          x = Deaths,                 # column with observed number of events
          n = Population,             # column with non-standard pops for each stratum
          stdpop = pop,               # standard populations for each stratum
          stdpoptype = "field")       # either "vector" for a standalone vector or "field" meaning std populations are in the data  

mortality_ds_rate_phe
```

### Indirectly standardized rates
```{r}

```

